Analysis methods for characterizing the climate sensitivity of annual growth and woody productivity using tree-rings

### 2020-05-12 doc/methods/analysis_methods.Rmd
update methods

1.2. Use the [climwin](https://cran.r-project.org/web/packages/climwin/index.html) R package [(van de Pol et. al. 2016)](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.12590), to identify the time window over which each candidate climate variable most strongly influences tree growth *for all species combined*. Specifically, we put the residuals in *climwin* with a mixed effects model using species and tree identity as random effect: `ln[residual] ~ [climate] +  + (1 | sp) + (1 | treeID)`. Outputs are saved [here](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/blob/master/results/ALL_species_mixed_model_on_residuals/). Example output is shown in Fig. 1.

@@

1.3. For each candidate climate variable, we move forward with that variable over the time window identified by climwin as a candidate climate variable for the multivariate models.

@@

2.1. Check colinearity among candidate climate variables; remove those that are redundant. (*Note: KAT thinks we need to reduce number of climate variables. ([GitHub issue # 14])(https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/14)*)

@@

![**Fig. 2. Current SCBI output for radial increment**. For each species, relationships are plotted if included in top model. Grey bars indicate long-term mean climate, shading indicates 1 SD. Time windows for each climate variable (CRU abbreviations) are indicated above plot. Note that we should probably reduce number of climate variables.](figures/ALL_variables_log_core_measurement.png)

![**Fig. 3. Current SCBI output for $\Delta AGB$**. For each species, relationships are plotted if included in top model. Grey bars indicate long-term mean climate, shading indicates 1 SD.  Time windows for each climate variable (CRU abbreviations) are indicated above plot. Note that we should probably reduce number of climate variables, and we are still working on the allometries. ](figures/ALL_variables_log_agb_inc.png)

>>>>

1.2. We use climate data from CRU v. 4.04. Based on how variables are related to one another, and how they are derived in CRU [Harris et al. 2020](https://rdcu.be/b3U3e), we grouped variables as follows, using climwin to eselct the best variable and time window for each: 
- Precip group (describes precip and its frequency): PRE, WET
- temperature group (describes temperature): TMP, TMIN, TMAX, PET
- temperature range group (describes variation in temperature, linked to cloud cover): DTR, CLD, PET

1.3. Use the [climwin](https://cran.r-project.org/web/packages/climwin/index.html) R package [(van de Pol et. al. 2016)](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.12590), to identify the climate variables and time windows that most strongly influences tree growth *for all species combined*. 
Climwin can check for linear and quadratic relationships. We checked quadratic only because (i) it speeds up the process, (ii) preliminarly tests revealled that quadratic was usually the one with lower AIC, and when not there tended to be little difference in AIC and the curve would look like a straight line anyway, (iii) quadratic makes more biological sense, (iv) it would be too tricky to code the gam to adapt to any circumstance.
We put the residuals derived in step 1.1 in *climwin* with a mixed effects model using species and tree identity as random effect: `ln[residual] ~ [climate] +  + (1 | sp) + (1 | treeID)`. 
Outputs are saved [here](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/blob/master/results/ALL_species_mixed_model_on_residuals/). Example output is shown in Fig. 1.

@@

1.3. For each group of candidate climate variables, we move forward with the best variable over the time window identified by climwin as a candidate climate variable for the multivariate models. When PET comes out as the best variable in both temperature and temperature range groups (should always be for the same time frame), then there are only 2 candidate variables for the GAM.

@@

2.1. The 2-3 candidate variables for the GAM should then go through the same process of checking for collinearity. If there is redundancy, we removed any variable with VIF > 10 (more standard way than looking at AIC).

@@

![**Fig. 2. OUTDATED Current SCBI output for radial increment**. For each species, relationships are plotted if included in top model. Grey bars indicate long-term mean climate, shading indicates 1 SD. Time windows for each climate variable (CRU abbreviations) are indicated above plot.](figures/ALL_variables_log_core_measurement.png)

![**Fig. 3. OUTDATED Current SCBI output for $\Delta AGB$**. For each species, relationships are plotted if included in top model. Grey bars indicate long-term mean climate, shading indicates 1 SD.  Time windows for each climate variable (CRU abbreviations) are indicated above plot. ](figures/ALL_variables_log_agb_inc.png)
